### 2025-09-18 talk/index.qmd
changed talk to dark theme

---
title: "Modelling potential environmental impacts of science activity in Antarctica"
title-slide-attributes:
    data-background-image: background-image.png
    data-background-opacity: "0.5"
author: 
  - name: David O'Sullivan
    affiliations: 
      - name: Geospatial Stuff
      - name: University of Auckland
  - name: Fraser Morgan
    affiliations: 
      - name: Bioeconomy Science Institute
format:
  revealjs:
    theme: black
    css: custom.scss
    transition: none
    navigation-mode: vertical
    fragment-in-url: true
    slide-number: false
    logo: logo.png
    footer: "Palmerston North Regional GIS Forum, September 2025"
    self-contained: true
---

## Overview {.smaller}
### [Study area](#study-area-1)
### [GPS data](#gps-data-1)
### [Building a hiking network](#building-a-hiking-network-1)
### [Results](#results-1)
### [Conclusions](#conclusions-1)

# Study area {.smaller background-image="background-image.png" background-opacity="0.5"}

##
![](fig-study-area-1.png)

# GPS data {.smaller background-image="background-image.png" background-opacity="0.5"}

##
::: {.columns}
::: {.column width=35%}
![](gps-device-picture.jpg)
:::
::: {.column width=65%}
Collected using QStarz Q1100P GPS Tracking Recorders between 2016 and 2018

40+ hour battery life and simple use

Location recorded every 30 seconds

Scientists turned devices on when leaving base camp in the morning and off on return at the end of the day
:::
:::

## The Dry Valleys {background-image="bing-maps-imagery.png"}
[Bing maps aerial imagery]{.aside}

## And on the ground {background-image="google-street-view.jpg"}
[Yup! There's streetview imagery]{.aside}

# Hiking networks {.smaller background-image="background-image.png" background-opacity="0.5"}

## Building a hiking network

Start with a regular lattice across the study area

Many topologies could be used

```{r}
library(sf)
library(igraph)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)

source("raster-to-graph-functions-paper.R")

theme_set(theme_void()) 
theme_update(
  plot.margin = unit(c(1, 1, 2, 1), "mm"),
  plot.title = element_text(hjust = 0.5, vjust = 2))

get_lattice <- function(dist, pts) {
  pts |>
    st_buffer(dist) |>
    st_contains(pts, sparse = FALSE) |>
    as.matrix() |>
    graph_from_adjacency_matrix(mode = "directed", diag = FALSE) |>
    set_vertex_attr("x", value = pts$x) |>
    set_vertex_attr("y", value = pts$y)
}

get_graph_plot <- function(G, main, crs = 2193) {
  return(ggplot() +
    geom_sf(data = G |> 
              get_graph_as_line_layer(crs = 2193), lwd = 0.35, colour = "white") +
    geom_sf(data = G |> 
              vertex_attr() |> as.data.frame() |> st_as_sf(coords = 1:2, crs = crs), 
            size = 8, fill = "black", colour = "white", shape = 21, lwd = 1) +
    coord_sf(xlim = c(-270, 270) + 1e6, 
             ylim = c(-270, 270) + 1e6, expand = FALSE) +
    ggtitle(main)) +
    theme(
      plot.background = element_rect(fill = "black"),
    )
}

grid_pts <- expand_grid(x = -20:20 * 100 + 1e6, y = -20:20 * 100 + 1e6) |>
  as.data.frame() |>
  st_as_sf(coords = c("x", "y"), crs = 2193, remove = FALSE)

hex_grid_pts <- grid_pts |>
  st_bbox() |> 
  st_make_grid(square = FALSE, what = "centers", cellsize = 107.457) |>
  st_as_sf() |>
  rename(geometry = x) |>
  mutate(geometry = geometry + c(-12, -3)) |>
  st_set_crs(2193)

xy <- hex_grid_pts |> st_coordinates()
hex_grid_pts <- hex_grid_pts |>
  mutate(x = xy[, 1], y = xy[, 2])

graphs <- mapply(
  get_lattice,
  # list(101, 150, 225, 110),
  list(101, 150, 110),
  # list(grid_pts, grid_pts, grid_pts, hex_grid_pts))
  list(grid_pts, grid_pts, hex_grid_pts))
# names(graphs) <- list("Rook's move", "Queen's move", "Knight's move", "Hex")
names(graphs) <- list("Rook's move", "Queen's move", "Hex")
```

```{r}
#| fig-width: 8
#| fig-height: 3.045
ggplots <- list()
for (n in names(graphs)) {
  ggplots[[n]] <- get_graph_plot(G = graphs[[n]], main = n)
}

wrap_plots(ggplots, nrow = 1) & theme(
  plot.background = element_rect(fill = "black"),
  text = element_text(colour = "white")
)
```

We went with hexagonal at ~100m spacing

## Up $\neq$ down {background-image="klondike.jpg" background-opacity="50%"}

Slope is critical to movement, so attach heights to nodes and calculate slope in each direction along lattice edges

Then apply a *hiking function* to estimate traversal times

::: aside
[People heading uphill in the Klondike gold rush 1897](https://www.nps.gov/klgo/learn/goldrush.htm) National Park Service, Klondike Gold Rush National Historical Park, Candy Waugaman Collection, KLGO Library SS-126-8831
:::

## Hiking functions
![](fig-hiking-functions-1.png)

## Smoothing GPS data
![](fig-gps-data-1.png)

## Curve fitting to cleaned data
![](fig-comparison-of-hiking-functions-1.png)

## Terrain-dependent _vs_ combined
![](fig-models-by-cover-compared-1.png)

## Completing and applying the network

Assign edges *estimated traversal times* based on slope, the hiking function, and land cover (moraine or rock)

Make the nodes and edges into a _directed graph_

Then use various _graph algorithms_ to find, e.g., everywhere-to-everywhere shortest paths

# Some outputs {.smaller background-image="background-image.png" background-opacity="0.5"}

##
![](fig-hiking-network-example-1.png)

## Betweenness centrality
After some exploration, we settled on _betweenness centrality_ as a graph metric that might be useful

This measure counts how many times each node appears on the shortest paths between every other pair of nodes

An indicator of the relative likelihood of each location being visited

##

![](fig-betweenness-map-1.png)

## Radius-limited betweenness
Restrict betweenness centrality to nodes no more than some time apart

This is _much_ faster to calculate (yay!)

Also... it looks like it has more value

It (perhaps) is relevant to how people navigate in such environments

##

![](fig-radius-limited-betweenness-1.png)

## Planning a path network?

One way to minimise impact might be to plan paths

This is _experimental_ at this stage

Based on a _minimum spanning tree_ approximation to an _arborescence_

::: aside
A minimum spanning tree is subgraph with minimum total cost that reaches all the nodes in a network. An arborescence is the equivalent in a directed graph.
:::

##

![](fig-betweenness-limited-sites-1.png)

##

![](fig-approx-minimum-arborescence-1.png)

---

## Conclusions
Terrain-differentiated data-fitted hiking functions are a novelty

Potential wider application of radius-limited betweenness centrality?

A manuscript almost finished

&nbsp;

#### For more
* On this project: [dosull.github.io/antarctica](https://dosull.github.io/antarctica)

* About Geospatial Stuff: [dosull.github.io](https://dosull.github.io)

# {background-image="background-image.png"}
[Questions?]{style="font-size:200px;color:black;opacity:25%;text-align:center;"}