### 2025-09-24 talk/index.qmd
many minor tweaks to slides

---
title: "Modelling potential environmental impacts of science activity in Antarctica"
title-slide-attributes:
    data-background-image: background-image.png
    data-background-opacity: "0.65"
author: 
  - name: David O'Sullivan
    affiliations: 
      - name: Geospatial Stuff
      - name: University of Auckland
  - name: Fraser Morgan
    affiliations: 
      - name: Bioeconomy Science Institute
format:
  revealjs:
    theme: [black, custom.scss]
    transition: none
    navigation-mode: vertical
    fragment-in-url: true
    slide-number: false
    logo: logo.png
    footer: "Palmerston North Regional GIS Forum, September 2025"
    self-contained: true
---

## Overview {.smaller background-image="miers-valley.jpg"}
### [Study area and data](#study-area-and-data-1)
### [Hiking networks](#hiking-networks-1)
### [Some outputs](#some-outputs-1)
### [Conclusions](#conclusions-1)

::: aside
Image: Miers Valley, from [commons.wikimedia.org](https://commons.wikimedia.org/wiki/File:Miers_Valley_CKL.jpg)
:::

# Study area and data {.smaller background-image="background-image.png" background-opacity="0.65"}

##
![](fig-study-area-1.png)

## GPS data
::: {.columns}
::: {.column width=35%}
![](gps-device-picture.jpg)
:::
::: {.column width=65%}
Collected using QStarz Q1100P GPS Tracking Recorders between 2016 and 2018

40+ hour battery life and simple use

Location recorded every 30 seconds

Scientists turned devices on when leaving base camp in the morning and off on return at the end of the day
:::
:::

## The Dry Valleys {background-image="bing-maps-imagery.png"}
[Bing maps aerial imagery]{.aside}

## And on the ground {background-image="google-street-view.jpg"}
[Yup! There's streetview imagery]{.aside}

# Hiking networks {.smaller background-image="background-image.png" background-opacity="0.65"}

## Building a hiking network
### Start with regular lattice across study area
### Many topologies could be used

```{r}
library(sf)
library(igraph)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)

source("raster-to-graph-functions-paper.R")

theme_set(theme_void()) 
theme_update(
  plot.margin = unit(c(1, 1, 2, 1), "mm"),
  plot.title = element_text(hjust = 0.5, vjust = 2))

get_lattice <- function(dist, pts) {
  pts |>
    st_buffer(dist) |>
    st_contains(pts, sparse = FALSE) |>
    as.matrix() |>
    graph_from_adjacency_matrix(mode = "directed", diag = FALSE) |>
    set_vertex_attr("x", value = pts$x) |>
    set_vertex_attr("y", value = pts$y)
}

get_graph_plot <- function(G, main, crs = 2193) {
  return(ggplot() +
    geom_sf(data = G |> 
              get_graph_as_line_layer(crs = 2193), lwd = 0.35, colour = "white") +
    geom_sf(data = G |> 
              vertex_attr() |> as.data.frame() |> st_as_sf(coords = 1:2, crs = crs), 
            size = 8, fill = "black", colour = "white", shape = 21, lwd = 1) +
    coord_sf(xlim = c(-270, 270) + 1e6, 
             ylim = c(-270, 270) + 1e6, expand = FALSE) +
    ggtitle(main)) +
    theme(
      plot.background = element_rect(fill = "black"),
    )
}

grid_pts <- expand_grid(x = -20:20 * 100 + 1e6, y = -20:20 * 100 + 1e6) |>
  as.data.frame() |>
  st_as_sf(coords = c("x", "y"), crs = 2193, remove = FALSE)

hex_grid_pts <- grid_pts |>
  st_bbox() |> 
  st_make_grid(square = FALSE, what = "centers", cellsize = 107.457) |>
  st_as_sf() |>
  rename(geometry = x) |>
  mutate(geometry = geometry + c(-12, -3)) |>
  st_set_crs(2193)

xy <- hex_grid_pts |> st_coordinates()
hex_grid_pts <- hex_grid_pts |>
  mutate(x = xy[, 1], y = xy[, 2])

graphs <- mapply(
  get_lattice,
  # list(101, 150, 225, 110),
  list(101, 150, 110),
  # list(grid_pts, grid_pts, grid_pts, hex_grid_pts))
  list(grid_pts, grid_pts, hex_grid_pts))
# names(graphs) <- list("Rook's move", "Queen's move", "Knight's move", "Hex")
names(graphs) <- list("Rook's move", "Queen's move", "Hex")
```

```{r}
#| fig-width: 8
#| fig-height: 3.045
ggplots <- list()
for (n in names(graphs)) {
  ggplots[[n]] <- get_graph_plot(G = graphs[[n]], main = n)
}

wrap_plots(ggplots, nrow = 1) & theme(
  plot.background = element_rect(fill = "black"),
  text = element_text(colour = "white")
)
```

### We went with hexagonal at ~100m spacing

## Up $\neq$ down {background-image="klondike.jpg" background-opacity="0.65"}
### Slope is critical to movement, so attach heights to nodes and calculate slope in each direction along lattice edges
### Then apply a *hiking function* to estimate traversal times

::: aside
[People heading uphill in the Klondike gold rush 1897](https://www.nps.gov/klgo/learn/goldrush.htm) National Park Service, Klondike Gold Rush National Historical Park, Candy Waugaman Collection, KLGO Library SS-126-8831
:::

## Hiking functions
![](fig-hiking-functions-1.png)

## Smoothing GPS data
![](fig-gps-data-1.png)

## Curve fitting to cleaned data
![](fig-comparison-of-hiking-functions-1.png)

## Terrain-dependent _vs_ combined
![](fig-models-by-cover-compared-1.png)

## Completing and applying the network
### Assign edges *estimated traversal times* based on slope, the hiking function, and land cover (moraine or rock)
### Make the nodes and edges into a _directed graph_
### Then use various _graph algorithms_ to find, e.g., everywhere-to-everywhere shortest paths

# Some outputs {.smaller background-image="background-image.png" background-opacity="0.65"}

##
![](fig-hiking-network-example-1.png)

## Betweenness centrality
### After some exploration, we landed on _betweenness centrality_ as a useful metric
### Counts how often each node appears on shortest paths between every other pair of nodes
&nbsp;

### &rarr; Indicator of relative likelihood of locations being visited

##

![](fig-betweenness-map-1.png)

## Radius-limited betweenness
### Restrict betweenness centrality to nodes no more than some cost (i.e., _time_) apart
### _Much_ faster to calculate
### Also... it may have more value: perhaps relevant to how people navigate in such environments

##

![](fig-radius-limited-betweenness-1.png)

## Planning a path network?
### A way to reduce impact might be to plan paths
### _Experimental_ at this stage
### Based on a _minimum spanning tree_ approximation to an _arborescence_

::: aside
A minimum spanning tree is a subgraph with minimum total cost that reaches all the nodes in a network. An arborescence is the equivalent in a directed graph.
:::

##

![](fig-betweenness-limited-sites-1.png)

##

![](fig-approx-minimum-arborescence-1.png)

---

## [Conclusions]{style="text-shadow:0px 0px 8px rgba(102,0,0,0.5);"} {background-image="beacon-valley.jpg" style="text-shadow:0px 0px 8px rgba(102,0,0,0.5);"}
### Terrain-differentiated data-fitted hiking functions are a novelty
### Potential wider application of radius-limited betweenness centrality?

&nbsp;

#### For more
* On this project: [[dosull.github.io/antarctica](https://dosull.github.io/antarctica)]{style="background-color:rgba(102,0,0,0.65);"}

* About ![](logo.png){style="vertical-align:middle;width:60px"} [Geospatial Stuff: [dosull.github.io](https://dosull.github.io)]{style="background-color:rgba(102,0,0,0.65);"}

::: aside
Image: Beacon Valley, from [reddit.com/rHumanForScale](https://www.reddit.com/r/HumanForScale/comments/by9q7f/beacon_valley_antarctica/)
:::

# {background-image="background-image.png"}
[Questions?]{style="font-size:200px;color:black;opacity:25%;text-align:center;"}